### 2019-03-14 distill_draft_notes/Notes_Brain_Storming_March/Notes_Brainstorming_March.tex
add notes about discussions this week

\documentclass[a4paper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{braket}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Physics of Equilibrium}
\author{Patrick, Juan-Miguel, Peter, Nathan}

\begin{document}
\maketitle

\section{Maximum Entropy Priciple}

\subsection{Theory}

If we maximize the entropy of the probabilitiy distribution $p(x)$ that describes the random variables $\{x_i\}_i^N$ under the constarints $\sum_i p(x_i) f_k(x_i) = \langle f_k \rangle$ we automatically obtain the Boltzmann distribution

\begin{equation*}
p(x) = 1/Z \sum_i^N \exp(\sum_k \lambda_k f_k(x_i)).
\end{equation*}

Where $f_k(x_i)$ are $m$ arbitrary functions of the random variables $x_i$ and $\lambda_k$ are the Lagrange multipliers . An additional constraint is that $\sum_i p(x_i) = 1$.

\subsection{Applied to physics}

In a physical system the constraints are given by the Hamiltonian of the system. For example through the expectation values of single site observables and two-body correlators.

\begin{align*}
	\langle \sigma_i \rangle &= \sum_i p(\sigma_i) \sigma_i \\
	\langle \sigma_i \sigma_j \rangle &= \sum_{i,j} p(\sigma_i, \sigma_j) \sigma_i \sigma_j 
\end{align*}

These constraints come directly from the Hamiltonian $H = \sum_i b_i \sigma_i  + \sum_{i,j} J_{i,j} \sigma_i \sigma_j$.

\textcolor{red}{Here are several things still unclear. Why does this meanfield approach apply? Are there other constarints we could implement?} One way to explain it is probably the sufficiency of the staistic. Which would mean that these two expectation values are sufficient statistics for the Boltzmann distribution. But this would mean, that the Boltzmann distribution comes also from the fact that we chose these statistics / constarints.

\subsubsection{Sufficient statistic}
A statistic $t(\vec{x})$ is sufficient with respect to a statistical model $p_{\theta}(\vec{x})$ and its parameter $\theta$ if  the conditional probability $p(\vec{x}|(t(\vec{x}))$ does not depend on $\theta$. Where $\vec{x}$ is a sample $(x_1, \dots x_n)$ of random variables. In other words, a statistic is sufficient with respect to a statistical model and the parameter $\theta$, if there is no way of extracting any additional information from the sample (concerning parameter $\theta$) than the staistic $t(\vec{x})$ already conveis.

A simple example: If we choose to describe a set of random variables by a Gaussian distribution $\mathcal{N}(\mu, \sigma^2)$, the mean of $\vec{x}$ is a sufficient statistic for the parameter $\mu$ .

\subsection{Connection to machine learning}
In ML we normally choose the model we would like to fit, which is in this case a Boltzmann distribution $p = \exp(\sum_k w_k f_k)$ and then we fit the data $(x_1, \dots x_n)$ such that $\langle \sigma_i \sigma_j \rangle_p = \langle \sigma_i \sigma_j \rangle_{\text{Data}}$ and $\langle \sigma_i \rangle_p = \langle \sigma_i  \rangle_{\text{Data}}$. To do so we apply several tricks like making BM bipartite and use sampling to get approximate the model distribution.

\subsection{Going from this theory to a trainable model}

\begin{itemize}
	\item We will start with a visualization of  $\langle \sigma_i \sigma_j \rangle_p = \langle \sigma_i \sigma_j \rangle_{\text{Data}}$ for a general energy based model where we don't have hidden and visible units. The input data just has the same dimension as the number of spins. The expectation values for the data are therefore fixed and we change the model parameters with sliders and the reader should try to reproduce the data distribution by changing the model parameters with these sliders.
	\item In a next step we can make a visualization where we do something similar but with a task that cannot be achieved. For example we have a 4 dimensional data set (4 pixels) and we want some configurations to be very likely and some very similar configurations have to be almost 0 probability. We will show that this cannot be achieved just by a system of 4 spins. Therefore we introduce hidden units which are just traced out in the end.
	\item With such a model (which is a Boltzmann machine) we actual have more expressive power but this power comes with the price that we have to sample the hidden units now also for the $\langle \sigma_i \sigma_j \rangle_{\text{Data}}$ part, which we now write as $\langle v_i h_j \rangle_{\text{Data}}$.
	\item We can now also add the same figure from step 1) again, but now the task is a bit trickier, because the target distribution $\langle \sigma_i \sigma_j \rangle_{\text{Data}} \rightarrow \langle v_i h_j \rangle_{\text{Data}}$ is not fixed anymore, because it also contains random variables $h_j$. Therefore it will be a bit harder to fit $\langle v_i h_j \rangle_p = \langle v_i h_j \rangle_{\text{Data}}$.
	\item from here on we can also explain for example why we do Restricted BMs. Why they are bipartite? Because of this the conditional probabilities $p(\vec{v} | \vec{h})$ and $p(\vec{h} | \vec{v})$ marginalize. And finally why contrastive divergence?
	\item \textcolor{red}{We might mention Hopfield networks in this section?}
\end{itemize}

\subsection{Equilibration with bath of data}

The formula $\langle \sigma_i \sigma_j \rangle_p = \langle \sigma_i  \sigma_j \rangle_{\text{Data}}$ made us think of this process as a thermalization with a bath of data, where the system has to equlibrate with this bath. But for each pair of spins we would need a different bath and the equilibrate temperature would be the parameter $W_{i,j}$. For the moment we decided that this picture does not help to understand BMs better. It is rather confusing and complicated.

\subsection{Contrastive Divergence}

The update rule of an RBM is:

\begin{equation*}
	\Delta W_{i,j} = \langle \frac{ \partial E}{\partial W_{i,j}} \rangle_{\text{Data}} - \langle \frac{\partial E}{\partial W_{i,j}} \rangle_{p_{\theta}} = \langle v_i h_j \rangle_{\text{Data}} - \langle v_i h_j \rangle_{p_{\theta}}
\end{equation*}

If the energy is given by $E(v,h) = \sum_{i,j} W_{i,j} v_i h_j $ (\textcolor{red}{sometimes there is a minus sign in front of the sum, which changes the direction}), then the update rule is $W \rightarrow W - \Delta W$

The question here was, why is the minus sign in front of the average over the model distribution and not vice versa. How do I know in which direction I will have to move my parameters, if I train a model.
There is a intuitive approach to this question if we consider the very extreme case of a system whith only two spins $v$ and $h$ and we assume that these variables are always not equal in the data and always equal if we sample them from the distribution $p_{\theta}$. This means that $\langle v_i h_j \rangle_{\text{Data}} = -1$
 and $\langle v_i h_j \rangle_{p_{\theta}}  = 1$. So the update will be minimally, namely $\Delta W_{i,j} = -2$. And this makes physically sense because the data is always anti-alligned, which means the minimum energy of the configurations of the data will be achieved by increasing the weights. Since the configurations of the distribution $p_{\theta}$ are always aligned to achieve minimum energy $W$ must be strongly negative.

\end{document}